---
title: "FTE Cleaning"
output:
  html_document:
    df_print: paged
---
Since we're using R Markdown files, you cannot change the working directory using setwd(). Using this function pulls the files from the working directory that is set here, not what is set as the working directory.
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data")
```

Only run this if the workspace needs to be cleared.
```{r}
##Clear working directory
rm(list=ls())

```

Load all of the libraries needed and set the working directory. The working directory is where the files are stored. When copying this code to a different computer, save the base files to a folder and change the working directory to that folder. That folder is where files will be read from and saved to.
```{r}

library(dplyr)
library(tidyr)
library(stringr)
library(base)
```

```{r}

opts_knit$set(root.dir = "/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data")
setwd("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data")
```

```{r}
master<-read.csv("/Users/miritfriedman/Desktop/Confirmed_v7.csv")

```


Read in the crosswalk file to link FIPS with GOV individual unit identifier codes.
```{r}
gov_fips<-read.csv("/Users/miritfriedman/Desktop/Squeeze Data/govID_to_FIPS.csv")
gov_fips<-gov_fips[c(2:4)]
```


```{r}
master<-data.frame(master, "ID", "YEAR",1997,2002,2007,2012,2017)
#create the tidy format with a year 1997-2018 for each FIPS code
master<-gather(master,"ID","YEAR","X1997","X2002","X2007","X2012","X2017",11:17)
master<-master[!(master$YEAR=="ID" | master$YEAR=="YEAR"),]
#master<-subset(master,select=-c(ID))
master<-subset(master,select=c(2,3,5,11))
master<-merge(master,gov_fips,by = "FIPS")
```


Complete the cleaning for the 2017 data.
```{r}

fte17<- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/17cempst.csv", header=FALSE)

#Separate the data into separate columns based on their position. 
fte17 <-
  tidyr::separate(
    data =fte17,
    col = V1,
    sep = c(2,3,6,9,12,14,17,20,30,31,32,44,45,46,56,57,58,70,71,72,82,83,84,94),
    into = c("State_Code", "Summary_Code", "County_Code", "Unit_ID","Supplemental_Code",
             "Sub_Code","B1",
             "Data_Function","Full Time Employees","B2",
             "FTE Data Flag","Full Time Payroll","B3",
             "FT Payroll Data Flag","Part Time Employees", "B4",
             "PT Data Flag", "Part Time Payroll", "B5",
             "Part Time Payroll Data Flag", "Part Time Hours","B6",
             "Part Time Hours Data Flag", "Full_Time_Equivalent_Employees"),
    remove = FALSE
  ) 

fte17 <-
  tidyr::separate(
    data =fte17,
    col = V1,
    sep = c(14),
    into = c("Individual Unit ID","Delete"),
    remove = TRUE
  ) 
fte17<-subset(fte17,select=-c(Delete))

fte17$State_Code<-as.numeric(fte17$State_Code)

fte17$ID<-paste(fte17$State_Code,fte17$Summary_Code, fte17$County_Code, fte17$Unit_ID, sep="")

#select out only the data totals for each place. The data function code for total is 000
fte17<-fte17[fte17$Data_Function == '000', ]


#Clip the file by the master list 
fte17<- filter(fte17,ID %in% master$GOVSid)
```

Complete the cleaning for the 2012 data
```{r}
fte12<- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/12cempst.csv", header=FALSE)

#Separate the data into separate columns based on their position. 
fte12 <-
  tidyr::separate(
    data =fte12,
    col = V1,
    sep = c(2,3,6,9,12,14,17,20,30,31,32,44,45,46,56,57,58,70,71,72,82,83,84,94),
    into = c("State_Code", "Summary_Code", "County_Code", "Unit_ID","Supplemental_Code",
             "Sub_Code","B1",
             "Data_Function","Full Time Employees","B2",
             "FTE Data Flag","Full Time Payroll","B3",
             "FT Payroll Data Flag","Part Time Employees", "B4",
             "PT Data Flag", "Part Time Payroll", "B5",
             "Part Time Payroll Data Flag", "Part Time Hours","B6",
             "Part Time Hours Data Flag", "Full_Time_Equivalent_Employees"),
    remove = FALSE
  ) 

fte12 <-
  tidyr::separate(
    data =fte12,
    col = V1,
    sep = c(14),
    into = c("Individual Unit ID","Delete"),
    remove = TRUE
  ) 
fte12<-subset(fte12,select=-c(Delete))

fte12$State_Code<-as.numeric(fte12$State_Code)

#create the complete ID code to be used.
fte12$ID<-paste(fte12$State_Code,fte12$Summary_Code, fte12$County_Code, fte12$Unit_ID, sep="")

#select out only the data totals for each place. The data function code for total is 000
fte12<-fte12[fte12$Data_Function == '000', ]

#Clip the file by the master list 
fte12<- filter(fte12,ID %in% master$GOVSid)
```


Complete the cleaning for the 2007 data.
```{r}
fte07<- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/07cempst.csv", header=FALSE)

#Separate the data into separate columns based on their position. 
fte07 <-
  tidyr::separate(
    data =fte07,
    col = V1,
    sep = c(2,3,6,9,12,14,17,20,30,31,32,44,45,46,56,57,58,70,71,72,82,83,84,94),
    into = c("State_Code", "Summary_Code", "County_Code", "Unit_ID","Supplemental_Code",
             "Sub_Code","B1",
             "Data_Function","Full Time Employees","B2",
             "FTE Data Flag","Full Time Payroll","B3",
             "FT Payroll Data Flag","Part Time Employees", "B4",
             "PT Data Flag", "Part Time Payroll", "B5",
             "Part Time Payroll Data Flag", "Part Time Hours","B6",
             "Part Time Hours Data Flag", "Full_Time_Equivalent_Employees"),
    remove = FALSE
  ) 

fte07 <-
  tidyr::separate(
    data =fte07,
    col = V1,
    sep = c(14),
    into = c("Individual Unit ID","Delete"),
    remove = TRUE
  ) 
fte07<-subset(fte07,select=-c(Delete))

fte07$State_Code<-as.numeric(fte07$State_Code)


#create the complete ID code to be used.
fte07$ID<-paste(fte07$State_Code,fte07$Summary_Code, fte07$County_Code, fte07$Unit_ID, sep="")

#select out only the data totals for each place. The data function code for total is 000
fte07<-fte07[fte07$Data_Function == '000', ]

#Clip the file by the master list 
fte07<- filter(fte07,ID %in% master$GOVSid)
```

Complete the cleaning for the 2002 data.
```{r}
fte02<- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/02cempst.csv", header=FALSE)

#Separate the data into separate columns based on their position. 
fte02 <-
  tidyr::separate(
    data =fte02,
    col = V1,
    sep = c(2,3,6,9,12,14,17,20,30,42,52,64,74,84),
    into = c("State_Code", "Summary_Code", "County_Code", "Unit_ID","Supplemental_Code",
             "Sub_Code","B1",
             "Data_Function","Full Time Employees","Full Time Payroll",
             "Part Time Employees", "Part Time Payroll", 
             "Part Time Hours","Full_Time_Equivalent_Employees"),
    remove = FALSE
  ) 

fte02 <-
  tidyr::separate(
    data =fte02,
    col = V1,
    sep = c(14),
    into = c("Individual Unit ID","Delete"),
    remove = TRUE
  ) 
fte02<-subset(fte02,select=-c(Delete))

fte02$State_Code<-as.numeric(fte02$State_Code)


#create the complete ID code to be used.
fte02$ID<-paste(fte02$State_Code,fte02$Summary_Code, fte02$County_Code, fte02$Unit_ID, sep="")

#select out only the data totals for each place. The data function code for total is 000
fte02<-fte02[fte02$Data_Function == '000', ]

#Clip the file by the master list 
fte02<- filter(fte02,ID %in% master$GOVSid)
```

Complete the cleaning for the 1997 data. 
```{r}
fte97<- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/97cempst.csv", header=FALSE)

#Separate the data into separate columns based on their position. 
fte97 <-
  tidyr::separate(
    data =fte97,
    col = V1,
    sep = c(2,3,6,9,12,14,17,20,30,42,52,64,74,84),
    into = c("State_Code", "Summary_Code", "County_Code", "Unit_ID","Supplemental_Code",
             "Sub_Code","B1",
             "Data_Function","Full Time Employees","Full Time Payroll",
             "Part Time Employees", "Part Time Payroll", 
             "Part Time Hours","Full_Time_Equivalent_Employees"),
    remove = FALSE
  ) 

fte97 <-
  tidyr::separate(
    data =fte97,
    col = V1,
    sep = c(14),
    into = c("Individual Unit ID","Delete"),
    remove = TRUE
  ) 
fte97<-subset(fte97,select=-c(Delete))

fte97$State_Code<-as.numeric(fte97$State_Code)

#create the complete ID code to be used.
fte97$ID<-paste(fte97$State_Code,fte97$Summary_Code, fte97$County_Code, fte97$Unit_ID, sep="")

fte97<-fte97[fte97$Data_Function == '000', ]
#Clip the file by the master list 
fte97<- filter(fte97,ID %in% master$GOVSid)
```

```{r}
fte97$YEAR<-1997
fte02$YEAR<-2002
fte07$YEAR<-2007
fte12$YEAR<-2012
fte17$YEAR<-2017

fte97<-subset(fte97, select=c(15:17))
fte02<-subset(fte02, select=c(15:17))
fte07<-subset(fte07, select=c(25:27))
fte12<-subset(fte12, select=c(25:27))
fte17<-subset(fte17, select=c(25:27))

fte_merge<-bind_rows(fte97,fte02,fte07,fte12,fte17)
fte_merge$Full_Time_Equivalent_Employees<-as.numeric(fte_merge$Full_Time_Equivalent_Employees)
fte_merge<-rename(fte_merge, "GOVSid"="ID")
fte_merge<-merge(fte_merge,gov_fips, by= "GOVSid")

write.csv(fte_merge, "/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/FTE Modified Files/final_fte_output.csv")

summary(fte_merge$Full_Time_Equivalent_Employees)
```


```{r}
#master$FTE <- ifelse(
    #is.na(match(paste(master$ID), paste(fte17$ID))),
    #fte17$Full_Time_Equivalent_Employees,
   # "NA")

master<-rename(master, "YEAR"="ID")
master<-separate(master, "YEAR", into = c("X" , "YEAR"), sep=1, remove = TRUE)
master<-subset(master, select = -c(4))

master$FTE<-
 with(master, if_else(master$YEAR == 2017 & master$GOVSid == fte17$ID, fte17$Full_Time_Equivalent_Employees, "NA"
  ))



master$FTE<-ifelse(master$YEAR == 2017 |master$GOVSid == fte17$ID, fte17$Full_Time_Equivalent_Employees, "NA")
                 

 ifelse(master$YEAR == 2012 |master$GOVSid == fte12$ID, fte12$Full_Time_Equivalent_Employees,
                         ifelse(master$YEAR == 2007 |master$GOVSid == fte07$ID, fte07$Full_Time_Equivalent_Employees,
                                ifelse(master$YEAR == 2002 |master$GOVSid == fte02$ID, fte02$Full_Time_Equivalent_Employees,
                                      ifelse(master$YEAR == 1997 |master$GOVSid == fte97$ID, fte97$Full_Time_Equivalent_Employees,"NA")))))
```


Merge the linking file to the FTE data 
```{r}
#ID_match_FTE<-merge(fte,gov_fips, by.x="ID",by.y="ID", no.dups = TRUE)

##filter out only the rows that have the data function of 000 which is the total - all government employment functions.
#ID_match_FTE<-filter(ID_match_FTE, Data_Function == "000")
#ID_match_FTE<-subset(ID_match_FTE, select=-c(Name))

```


Subset out the columns of interest
```{r}
fte_output<-subset(fte_output, select=c(FIPS, ID, PWSID, Full_Time_Equivalent_Employees, PWS.Name))
```

Save File Output
```{r}
write.csv(master, "/Users/miritfriedman/Desktop/Squeeze Data/fte/FTE data/FTE Modified Files/final_fte_output.csv")
```

Descriptive Statistics

